---
title: "STAT 425 Assignment 2"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Name: Kimmy Liu
### Netid: zl32

Submit your computational work both as an R markdown (*.Rmd) document and as a pdf, along with any files needed to run the code. Embed your answers to each problem in the document below after the question statement. If you have hand-written work, please scan or take pictures of it and include in a pdf file, ideally combined with your pdf output file from R Markdown.

## Problem 3
We aim to minimize the function 
$$
f(\beta)=\sum_{i=1}^{n}\left[y_{i}-\beta x_{i}\right]^{2}
$$
Take the first order derivative of $f(\beta)$, we get
$$
\begin{aligned}
f^{\prime}(\beta)&=\sum_{i=1}^{n} 2\left(y_{i}-\beta x_{i}\right)\left(-x_{i}\right)\\\
&=-2 \sum_{i=1}^{n} x_{i} y_{i}+2 \beta \sum_{i=1}^{n} x_{i}^{2}
\end{aligned}
$$
And the second order derivative is 
$$
f^{\prime\prime}(\beta)= 2 \sum_{i=1}^{n} x_{i}^{2} \geq 0
$$
Which indicates this function is a convex function. We can then find the optimal $\hat{\beta}$ by setting $f^{\prime}(\beta)=0$. Finally we get 
$$
\hat{\beta}=\sum_{i=1}^{n} x_{i} y_{i} / \sum_{i=1}^{n} x_{i}^{2}
$$

## Problem 4
**a)**

**Answer:**
$$
\begin{aligned}
\hat{\beta}&=\sum_{i=1}^{n} x_{i} y_{i} / \sum_{i=1}^{n} x_{i}^{2}\\
&=2.5
\end{aligned}
$$

**b)**

**Answer:**
The estimate $\hat{\beta}$ is

```{r}
x <- c(1,3,4,5,6,7,8)
y <- c(3,6,12,13,17,16,19)
fit <- lm(y ~ x+0)
fit$coef
```
Now we create a scatterplot.

```{r}
plot(x, y, xlim=c(0,8), ylim=c(0,20),
    main="Scatterplot",
    xlab="x",
    ylab="y")
abline(a=0,b=fit$coeff)
```

**c)**

**Answer:**
We have $e=y-\hat{y}=y-\hat{\beta}x$
```{r}
e <- y - fit$coeff *x
res<- data.frame(x=x,y=y,e=e)
knitr::kable(res,format = "markdown")
```

And the sumation is 
```{r}
sum(e)
```
Which is not zero. Since in this model there's no intercept and the vector of all 1’s, e does not have to be orthogonal to it. Thus the sumation does not have to be zero.












```{r}
x <- c(1.4,1.6,1.6,2,2,2.4,2.4,2.6)
y <- c(201,169,209,185,185,241,201,249)
fit <- lm(y ~ x)
summary(fit)
fit
```


```{r}


### (a)
SXX = sum( (x-mean(x))^2 )
SXY = sum( (x-mean(x)) * (y-mean(y)) )
beta1.hat = SXY/SXX
beta0.hat = mean(y) - beta1.hat * mean(x) 
e = y - (beta0.hat + beta1.hat * x)
n = length(y)
s2_e <- sum(e^2)/(n-2)
SXX
SXY
s2_e

```
```{r}
anova(fit)
```





```{r}
x = c(4,6,6,9,12,4,4,8,9,8)
y = c(254,330,310,384,398,350,390,442,480,422)

### (a)
SXX = sum( (x-mean(x))^2 )
SXY = sum( (x-mean(x)) * (y-mean(y)) )
beta1.hat = SXY/SXX
beta0.hat = mean(y) - beta1.hat * mean(x) 
e = y - (beta0.hat + beta1.hat * x)
n = length(y)
s2_e <- sum(e^2)/(n-2)
beta1.hat
beta0.hat
e
s2_e
sum(e^2)
n
### (b)
#2 * (1 - pt( abs( beta1.hat/sqrt(s2_e/SXX ) ), n-2 ))

# use R function
#summary(fit)
#summary(fit)$coefficients[2, 4]

### (c)
#1 -  pt( (beta1.hat - 4)/sqrt(s2_e/SXX ), n-2 )
```











