---
title: "STAT 420 Assignment 3"
output:
  html_document:
    df_print: paged
  pdf_document: default
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Name: Kimmy $\text{\underline{Liu}}$
### Netid: zl32@illinois.edu

## Problem 1


**a)** 

**Answer:**
```{r}
x <- c(2,2,2,3,2,2,3,4,3,4,3,4,3,4,4,3)
y <- c(420,570,590,610,660,780,890,930,950,1010,1050,1060,1090,1100,1200,1186)
SXX = sum((x-mean(x))**2)
SXY = sum((x-mean(x))*(y-mean(y)))
SYY = sum((y-mean(y))**2)
beta_hat1 = SXY/SXX
beta_hat0 = mean(y) - beta_hat1*mean(x)
print(beta_hat0)
print(beta_hat1)
```
we can also ues:
```{r}
fit <- lm(y~x)
fit
```

As a result, y_hat = 197+228x

**b)** 

**Answer:**

```{r}
y_hat = beta_hat0+beta_hat1*x
e <- y - y_hat
res<- data.frame(x=x,y=y,e=e)
knitr::kable(res,format = "markdown")
```
```{r}
sum(e)
```
Yes, the sum of the residuals equal zero. 

**c)** 

**Answer:**

```{r}
summary(fit)$sigma^2
```
**d)** 

**Answer:**
```{r}
summary(fit)$r.squared
```
about 57.93251% can be explained by a straight-line relationship. 
**e)** 

**Answer:**
```{r}
plot(x, y, main="Scatterplot",
xlab="bedroom number(x)",
ylab="monthly rent(y)")
abline(fit$coefficients)
```

## Problem 2


**a)** 

**Answer:**

use last 4 digits of my UIN:

establish the model:
```{r}
x = runif(n = 11, 10, 20)
beta_0 = 10
beta_1 = 8
#667755183
last4digit = 5183
set.seed(last4digit) 
sim_slr = function(x,
                   beta_0 = 10,
                   beta_1 = 8,
                   sigma = 6) {
  n = length(x)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  data.frame(predictor = x, response = y)
}
set.seed(last4digit)
sim_data = sim_slr(x,
                   beta_0 = 10,
                   beta_1 = 8,
                   sigma = 6)
```

simulate this model:
```{r}
sim_fit = lm(response ~ predictor, data = sim_data)
sim_fit$coefficients
```
plot:
```{r}
plot(
  response ~ predictor,
  data = sim_data,
  xlab = "Simulated Predictor Variable",
  ylab = "Simulated Response Variable",
  main = "Simulated Data",
  pch  = 20,
  cex  = 2,
  col  = "navyblue"
)
abline(sim_fit,
       lwd = 2,
       lty = 1,
       col = "darkorange",
       type = "solid")
abline(beta_0,
       beta_1,
       lwd = 1,
       lty = 2,
       col = "dodgerblue",
       type = "dashed")
legend(
  "topright",
  c("Estimate", "Truth"),
  lty = c(1, 2),
  lwd = c(2, 1),
  col = c("darkorange", "dodgerblue")
)
```


**b)** 

**Answer:**
```{r}

```

```{r}

plot(
  x , 10+8*x, type = "n",
  data = sim_data,
  xlab = "Simulated Predictor Variable",
  ylab = "Simulated Response Variable",
  main = "Simulated Data",
  pch  = 20,
  cex  = 2,
  col  = "navyblue"
)
beta_hat_1 = rep(0, 200)
beta_hat_0 = rep(0, 200)
for(i in 1:200) {
  sim_data = sim_slr( x = x, beta_0 = 10, beta_1 = 8, sigma = 6)
  model = lm(response ~ predictor, data = sim_data)
  beta_hat_1[i] = coef(model)[2]
  beta_hat_0[i] = coef(model)[1]
  abline(beta_hat_0[i],beta_hat_1[i],
       lwd = 2,
       lty = 1,
       col = "darkorange",
       type = "solid")
}

abline(beta_0,
       beta_1,
       lwd = 3,
       lty = 2,
       col = "white",
       type = "solid")

legend(
  "topright",
  c("Estimate", "Truth"),
  lty = c(1, 2),
  lwd = c(2, 3),
  col = c("darkorange", "white")
)
```




## Problem 4
**a)**

**Answer:**
Using the answer from the problem three, the answer is 2.5

**b)**

**Answer:**
The estimate value is

```{r}
x <- c(1,3,4,5,6,7,8)
y <- c(3,6,12,13,17,16,19)
fit <- lm(y ~ x+0)
fit$coef
```
Now we create a scatterplot.

```{r}
plot(x, y, xlim=c(0,8), ylim=c(0,20),
    main="Scatterplot",
    xlab="x",
    ylab="y")
abline(a=0,b=fit$coeff)
```

**c)**

**Answer:**
We have 
```{r}
e <- y - fit$coeff *x
res<- data.frame(x=x,y=y,e=e)
knitr::kable(res,format = "markdown")
```

And the sumation is 
```{r}
sum(e)
```
Which is not zero. Since in this model there's no intercept and the vector of all 1’s, e does not have to be orthogonal to it. Thus the sumation does not have to be zero.

```{r}
x1 = c(4,6,6,9,12,4,4,8,9,8)
x2 = c(12,12,12,12,12,18,18,18,18,18)

y = c(254,330,310,384,398,350,390,442,480,422)
fit = lm(y ~ x1+x2)
summary(fit)

```

```{r}
new = data.frame(x1=10, x2=10)
predict.lm(fit, new, interval = c("confidence"), level = 0.95)
```
```{r}
R2 = 1 - RSS/SYY
adj=1-(n-1)/(n-p)*(1-R2)
```

```{r}
library(faraway)
?sat
```


